From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-12 WOLFHPC15/article/evaluation.tex
fix evaluation helene 08_08
new evaluation helene 12_08

Navier-Stokes equations is a well known set of PDEs in fluid dynamics to simulate a flow evolution in time. At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}). Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{HeleneLS13}.

@@

The series-parallel tree decomposition $\Gamma_{tsp}$ of this simulation, extracted by the MSC transformation, is composed of seventeen sequence nodes and eighteen parallel nodes. Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks to perform concurrently, the number of time this level is observed in the final component assembly. One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1). As a level of sixteen parallel tasks is reached too times, and also five times for the level six, sequential restrictions could be amortized. However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching at its limits, but not to directly get performance. Moreover, as the level of parallelism in the application is heterogenous, the number of threads to launch for task parallelism, and the number of cores to keep for data parallelism is difficult to choose.

@@

Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds. Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb DDR3. 
One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph. Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear. The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
The fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations. Actually, a \emph{N-shape} could perfectly be handled by a dynamic scheduler for example. It has to be added, however, that the computation of $\Gamma_{tsp}$ is still interesting to detect how space loops can be fusionned in $\Gamma_{data}$.

@@

Evaluations have been performed on TGCC Curie \emph{thin nodes}. Each node is equiped with an Intel E5-2680 (Sandy-Bridge EP) processor containing 16 cores, 64GB of memory, and is linked to an Infiniband network. 
\fix{HC: what are the experiments presented ? Size of the mesh, number of time iterations}

Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores. Figure~\ref{fig:speedup} illustrates a speedup comparison, using the minimum sequential reference time (which is the MSL sequential time). Finally, Figure~\ref{fig:log2} illustrates the logarithmic scale of execution times as a function of the logarithmic scale of the number of cores.
\begin{figure}[!h]
 \begin{center}
 \begin{tabular}{c|c|c|c|c|c|c}
   Cores & 8 & 16 & 32 & 64 & 128 & 256\\
   \hline
   MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
   \hline
   SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
 \end{tabular}
\caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
\label{fig:times}
 \end{center}
\end{figure}
\begin{figure*}
\begin{center}
\subfloat[\label{fig:speedup}]{
\resizebox{8cm}{!}{\includegraphics{./images/speedup.pdf}}
}
\hspace{10pt}
\subfloat[\label{fig:log2}]{
\resizebox{8cm}{!}{\includegraphics{./images/logtime.pdf}}
}
\end{center}
\caption{(a) Speedup comparison, and (b) logarithmic execution times, both for the Shallow water equations, using pure SkelGIS and MSL with the SkelGIS data structure.}
\label{fig:perfs}
\end{figure*}

>>>>

Navier-Stokes equations is a well known set of PDEs in fluid dynamics to simulate a flow evolution in time. At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}). Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}. The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil kernels and sixty-six local computations.

@@

The series-parallel tree decomposition $\Gamma_{tsp}$ of this simulation, extracted by the MSC transformation, is composed of seventeen sequence nodes and eighteen parallel nodes. Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks to perform concurrently, the number of time this level is observed in the final component assembly. One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1). As a level of sixteen parallel tasks is reached too times, and also five times for the level four, sequential restrictions could be amortized. However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching its limits, but not to use alone. Moreover, as the level of parallelism in the application is heterogenous, the number of threads to launch for task parallelism, and the number of cores to keep for data parallelism is difficult to choose.

@@

Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds. Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb DDR3. 
One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph. Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear. The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections. However, a \emph{N-shape} could perfectly be handled by a dynamic scheduler, for example.
Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations. It has to be notice, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be fusionned in $\Gamma_{data}$.

@@

Evaluations have been performed on two clusters of \emph{Grid'5000}~\footnote{\url{https://www.grid5000.fr/mediawiki/index.php/Grid5000:Home}}. Figure~\ref{tab:hard} shows hardware configuration of tohse clusters. On the cluster \emph{Edel} (Grenoble site) have been performed three different experiments, each one described with the size of the Cartesian mesh, and the number of time iterations, as follows: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations. On the cluster \emph{Paravence} (Rennes sites), the two last experiments have been performed.

\begin{figure}[!h]
\begin{center}
 \begin{tabular}{|c|c|c|}
\hline
    Cluster & \textbf{Edel} & \textbf{Paravance}\\
     \hline         
    Proc. & 2 Intel Xeon E5520 & 2 Intel Xeon E52630v3\\
    & (2.27 GHz) & (2.4 GHz)\\
    Cores/n & 8 & 8\\
    Mem./n & 24 GB & 128 GB\\
    Compiler & OpenMPI & OpenMPI\\
    Network & InfiniBand 40G & 10 Gigabit Ethernet\\
\hline
 \end{tabular}
   \caption{\label{tab:grappe}Hardware of clusters used in the experiments.}
 \end{center}
\end{figure}

%Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores. 
Figure~\ref{fig:speedup} illustrates a speedup comparison, using the minimum sequential reference time (which is the MSL sequential time). Finally, Figure~\ref{fig:log2} illustrates the logarithmic scale of execution times as a function of the logarithmic scale of the number of cores.
% \begin{figure}[!h]
%  \begin{center}
%  \begin{tabular}{c|c|c|c|c|c|c}
%    Cores & 8 & 16 & 32 & 64 & 128 & 256\\
%    \hline
%    MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
%    \hline
%    SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
%  \end{tabular}
% \caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
% \label{fig:times}
%  \end{center}
% \end{figure}
\begin{figure*}
\begin{center}
\subfloat[Edel\label{fig:speedup1}]{
\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
}
\hspace{5pt}
\subfloat[Edel\label{fig:log21}]{
\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
}
\hspace{5pt}
\subfloat[Paravance\label{fig:speedup2}]{
\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
}
\hspace{5pt}
\subfloat[Paravance\label{fig:log22}]{
\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
}

\subfloat[Edel\label{fig:speedup3}]{
\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
}
\hspace{5pt}
\subfloat[Edel\label{fig:log23}]{
\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
}
\hspace{5pt}
\subfloat[Paravance\label{fig:speedup4}]{
\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
}
\hspace{5pt}
\subfloat[Paravance\label{fig:log24}]{
\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
}
\end{center}
\caption{(a) Speedup comparison, and (b) logarithmic execution times, both for the Shallow water equations, using pure SkelGIS and MSL with the SkelGIS data structure.}
\label{fig:perfs}
\end{figure*}
